### 2017-09-16 paper/DPC_T&A.tex
paper first commit

\documentclass[review]{elsarticle}

\usepackage{lineno,hyperref}
\modulolinenumbers[5]

\journal{Journal of \LaTeX\ Templates}

%%%%%%%%%%%%%%%%%%%%%%%
%% `Elsevier LaTeX' style
\bibliographystyle{elsarticle-num}
%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\begin{frontmatter}

\title{Data Predictive Control: Theory and Applications\tnoteref{mytitlenote}}
\tnotetext[mytitlenote]{The research leading to these results has received funding from the Italian Government under Cipe resolution n.135 (Dec. 21, 2012), project \emph{INnovating City Planning through Information and Communication Technologies} (INCIPICT).}

\author[FSaddress,EC]{Francesco Smarra\corref{mycorrespondingauthor}}
\cortext[mycorrespondingauthor]{Corresponding author}
\ead{francesco.smarra@univaq.it@univaq.it}

\author[AJaddress,EC]{Achin Jain}
\ead{achinj@seas.unpenn.edu}
\author[TDRaddress,EC]{Tullio De Rubeis}
\ead{tullio.derubeis@graduate.univaq.it}
\author[TDRaddress]{Dario Ambrosini}
\ead{dario.ambrosini@ing.univaq.it}
\author[FSaddress]{Alessandro D'Innocenzo}
\ead{alessandro.dinnocenzo@univaq.it}
\author[AJaddress]{Rahul Mangharam}
\ead{rahulm@seas.upenn.edu}

\address[EC]{Equal contribution}
\address[FSaddress]{Department of Information Engineering, Computer Science and Mathematics, Università degli Studi dell’Aquila, L’Aquila, Italy}
\address[TDRaddress]{Department of Industrial and Information Engineering and Economics, Università degli Studi dell’Aquila, L’Aquila, Italy}
\address[AJaddress]{Department of Electrical and Systems Engineering, University of Pennsylvania, Philadelphia, USA}

\begin{abstract}
TBD
\end{abstract}

\begin{keyword}
Energy efficiency, predictive control, machine learning.
\end{keyword}

\end{frontmatter}

%\linenumbers

\section{Introduction}

\paragraph{Motivations} TBD

\paragraph{State of the art} TBD

\paragraph{Contributions} TBD

\paragraph{Paper organization} TBD

\section{Problem definition and objectives} TBD

\section{Data Predictive Control using Random Forests} TBD

\subsection{Modeling} TBD

\subsection{Predictive Control} TBD

\section{Case study 1: bilinear model} TBD

\section{Case study 2: EnergyPlus model} TBD

\section{Case study 3: real house} TBD

\subsection{Description of the house} TBD

\subsection{EnergyPlus model} TBD

\subsection{Simulation Results} TBD

\paragraph{Energy-efficiency improvements}

\section{Conclusion and future work} TBD

\section*{References}

\bibliography{mybibfile}

\end{document}