### 2017-09-17 paper/abstract.tex
dumped CDC paper

Decisions on how to best operate large complex plants such as natural gas processing, oil refineries, and energy efficient buildings are becoming ever so complex that model-based predictive control (MPC) algorithms must play an important role. However, a key factor prohibiting the widespread adoption of MPC, is the cost, time, and effort associated with learning first-principles dynamical models of the underlying physical system. An alternative approach is to employ learning algorithms to build black-box models which rely only on real-time data from the sensors. Machine learning is widely used for regression and classification, but thus far data-driven models have not been used for closed-loop control. We present novel Data Predictive Control (DPC) algorithms that use Regression Trees and Random Forests for receding horizon control. 
%DPC also utilizes Gaussian Processes for smart experiment design by recommending how the accuracy of the current black-box models can be improved. 
We demonstrate the strength of our approach with a case study on a bilinear building model identified using real weather data and sensor measurements. 
In a one-to-one comparison, we show that DPC explains 70\% variation in the MPC controller.
%We validate our results by a one-to-one comparison against a benchmark MPC controller. 
We further apply DPC to a large scale multi-story EnergyPlus building model to curtail total power consumption in a Demand Response setting. 
In such cases, when the model-based controllers fail due to modeling cost, complexity and scalability, our results show that DPC curtails the desired power usage with high confidence.