### 2017-09-29 paper/main.tex

\documentclass[review]{elsarticle}

\usepackage{achinDefs}
\usepackage{lineno,hyperref}
\usepackage{slashbox}
\modulolinenumbers[5]
\usepackage{gensymb}
\usepackage{multirow}
\usepackage{eurosym}

\journal{Journal of \LaTeX\ Templates}

\begin{document}

\begin{frontmatter}

\title{Data Predictive Control with Random Forests:\\Theory and Applications\tnoteref{mytitlenote}}
\tnotetext[mytitlenote]{The research leading to these results has received funding from the Italian Government under Cipe resolution n.135 (Dec. 21, 2012), project \emph{INnovating City Planning through Information and Communication Technologies} (INCIPICT).}

\author[FSaddress,EC]{Francesco Smarra\corref{mycorrespondingauthor}}
\cortext[mycorrespondingauthor]{Corresponding author}
\ead{francesco.smarra@univaq.it}

\author[AJaddress,EC]{Achin Jain}
\ead{achinj@seas.unpenn.edu}
\author[TDRaddress,EC]{Tullio de Rubeis}
\ead{tullio.derubeis@graduate.univaq.it}
\author[TDRaddress]{Dario Ambrosini}
\ead{dario.ambrosini@univaq.it}
\author[FSaddress]{Alessandro D'Innocenzo}
\ead{alessandro.dinnocenzo@univaq.it}
\author[AJaddress]{Rahul Mangharam}
\ead{rahulm@seas.upenn.edu}

\fntext[EC]{Equal contribution}

\address[FSaddress]{Department of Information Engineering, Computer Science and Mathematics, Universit\`{a} degli Studi dell'Aquila, L'Aquila, Italy}
\address[AJaddress]{Department of Electrical and Systems Engineering, University of Pennsylvania, Philadelphia, USA}
\address[TDRaddress]{"G. Parolini Lab" - Department of Industrial and Information Engineering and Economics, Universit\`{a} degli Studi dell'Aquila, L'Aquila, Italy}

%% ABSTRACT
\begin{abstract}
\input{abstract.tex}
\end{abstract}

\begin{keyword}
Energy efficiency, demand response, predictive control, machine learning, EnergyPlus modeling, real application
\end{keyword}

\end{frontmatter}

%% SECTIONS
\input{intro.tex}

\input{dpc.tex}

\input{casestudyBilinear.tex}

\input{casestudyEPlus.tex}

\input{casestudyReal.tex}

%% CONCLUSION
\input{conclusion.tex}

%ACKNOWLEDGMENTS
\input{ack.tex}

%% BIBLIOGRAPHY
\section*{References}
\bibliographystyle{elsarticle-num}
\bibliography{fsRefs,achinRefs}

\end{document}