### 2017-09-30 paper/abstract.tex
some changes

Model-based control techniques have been widely used over the past years in control systems to improve system performance of simple controllers, such as PID, bang-bang, etc. Among these, one of the most popular is Model Predictive Control (MPC), due to its capacity to optimally control a system, based on the capability to predict its behavior over a future horizon. A key factor prohibiting the widespread adoption of MPC for complex systems is related to the difficulties (cost, time, and effort) associated with the model identification. Due to their lack of dynamical equations to predict real systems behavior, data-driven models are not suitable for control. We introduce a novel idea for predictive control based on system data, Data Predictive Control (DPC), leveraging machine learning algorithms, in particular regression trees and random forests. We show, using a bilinear model, that DPC has comparable performance with MPC. We further apply DPC to the Demand Response problem of a large scale multi-story EnergyPlus building model, and show that DPC curtails the desired power usage with high confidence. Finally, we use DPC to optimize ON/OFF scheduling of the heating system of a real house in order to minimize the power consumption while guaranteeing thermal comfort for occupants. Our results show that with DPC we obtain an energy saving in terms of primary energy up to $49.2\%$ when compared to a bang-bang controller.