### 2017-09-30 paper/abstract.tex
final edits from achin

Model-based control techniques have been widely used over the past years in control systems to improve system performance of simple controllers, such as PID, bang-bang, etc. Among these, one of the most popular is Model Predictive Control (MPC) due to its ability to predict the system's behavior over a future horizon and handle constraints in the optimal control setup. A key factor prohibiting the widespread adoption of MPC for complex systems is related to the difficulties (cost, time, and effort) associated with the model identification. 
On the other hand, data-driven models are not suitable for control because of non-physical models.
%Due to their lack of dynamical equations to predict real systems behavior, data-driven models are not suitable for control. 
We introduce a novel idea for predictive control based on system data, Data Predictive Control (DPC), leveraging machine learning algorithms, in particular regression trees and random forests. Using a bilinear model, we show that DPC has comparable performance with MPC. We further apply DPC to the Demand Response problem of a large scale multi-story EnergyPlus building building model, and show that DPC curtails the desired power usage with high confidence. Finally, we use DPC to optimize ON/OFF scheduling of the heating system of a real house in order to minimize the power consumption while guaranteeing thermal comfort for occupants. Our results show that with DPC we obtain an energy saving in terms of primary energy up to $49.2\%$ when compared to a bang-bang controller.