### 2017-12-24 paper/final/dpc.tex

\section{DATA PREDICTIVE CONTROL}
\label{S:dpc}
TO IMPROVE $\rightarrow$ \textcolor[rgb]{0,0,1}{Achin and Francesco}
\begin{itemize}
	\item Add more details
	\item Review notation
	\item Improve procedure and algorithm description
	\item Add optimization setup discussion
\end{itemize}

%The central idea behind DPC is to obtain control-oriented models using machine learning or black-box modeling, and formulate the control problem in a way that receding horizon control (RHC) can still be applied and the optimization problem can be solved efficiently.
%
%Consider a black-box model given by $x_{k+1}=f(x_k,u_k,d_k)$, where $x,u,d$ represent states, inputs and disturbances, respectively. Depending upon the learning algorithm, $f$ is typically nonlinear, nonconvex and sometimes nondifferentiable (as is the case with regression trees and random forests) with no closed-form expression. 
%Such functional representations learned through black-box modeling may not be directly suitable for control and optimization as the optimization problem can be computationally intractable, or due to nondifferentiabilities we may have to settle with a sub-optimal solution using evolutionary algorithms \cite{Kusiak2009}.
%These problems can be eliminated by decomposing 
%\begin{align}
%f(x_k,u_k,d_k)= g(x_k,h(u_k),d_k),
%\label{E:sepvars}
%\end{align}
%where both $g$ and $h$ are learned using the data, and $h(u_k)$ is convex and differentiable, and thus suitable for optimization. DPC uses this functional decomposition or \textit{separation of variables} to overcome the aforementioned challenges with black-box optimization.
%
%\subsection{Separation of Variables}
%\label{SS:sepvar}
%
%We distinguish between two sets of variables: control (or manipulated) variables $\tX^c \in \R^c$ and disturbance (or non-manipulated) variables $\tX^d \in \R^d$. The union of the two sets forms the full feature set for training, i.e. $\tX \equiv \tX^c \cup \tX^d \in \R^{c+d}$.
%Our goal is to replace a model-based controller with a data-driven controller, where the latter depends only on the historical sensor data. 
%These measurements could directly represent one or more states in the model-based control framework. We denote these as outputs $\tY \in \R$ for training, i.e. a $\tY$ represents a particular output and we can have separate models for multiple outputs. We define the number of training samples by $|(\tX,\tY)| = n$.
%
%Using separation of variables, the training process is divided into two steps. \textit{Step 1:} The trees and the ensembles are trained only on $\tX^d$, which eases the computational complexity. It is important to note that besides external disturbances, $\tX^d$ also contains autoregressive terms of the output $\tY$ which is the main reason for the state space explosion. \textit{Step 2:} Linear regression models are trained in the leaves (or terminal nodes) of the trees which are function of only $\tX^c$. We have validated this linear model assumption in \cite{Jain2016}. As we shall see in Section~\ref{SS:dpcrt} and \ref{SS:dpcrf}, the second step reduces the run-time control problem into a convex program. This process is illustrated in Figure~\ref{F:dpc-sepvars}.
%
%\subsection{DPC-RT: DPC with Regression Trees}
%\label{SS:dpcrt}
%When the data has lots of features, which interact in complicated, nonlinear ways, assembling a single global model such as linear or polynomial regression can be difficult, and can lead to poor response predictions.
%An approach to non-linear regression is to partition the data space into smaller regions, where the interactions are more manageable. 
%This partition is repeated recursively until finally we get to small chunks of the data space where we can fit simple (eg. linear parametric) models. 
%Therefore, in \eqref{E:sepvars}, the global model $f$ has two parts: the recursive partition $g$, and a linear (and convex) model $h$ for each cell of the partition.
%
%\begin{figure}[t!]
%	\centering
%	\includegraphics[width=20pc]{figures/dpc-sepvars.eps}
%	\caption{Separation of variables. \textit{Step 1:} Tree $\mathcal{T}_1$ is trained only on the disturbances $\tX_d$ as the features. Tree $\mathcal{T}_2$ uses both the disturbances $\tX_d$ and the control variables $\tX_c$ for splitting and is thus not computationally suitable for control. \textit{Step 2:} In the leaf $R_i$ of the trees, a linear regression model parametrized by $\beta_i$ is defined as a function only of the control variables.}
%	\captionsetup{justification=centering}
%	\label{F:dpc-sepvars}
%\end{figure}
%
%Now, our goal is to predict the state $\tY$ at time $k$ for next $N$ time steps, i.e. $\tY_{\mathrm{k+1|k}},\dots,\tY_{\mathrm{k+N|k}}$, where $N$ is the control horizon. Applying the separation of variables, we build $N$ regression trees using CART procedure \cite{Breiman1984} such that the output $\tY_{\mathrm{k+j|k}}$ of the $j^{th}$ tree depends upon the previous $N$ disturbances:
%\begin{gather}
%\label{E:model_tree}
%\tY_{\mathrm{k+j|k}} = \mathit{f}_{\mathrm{tree}} \left( \tX^d_{\mathrm{k+j-N|k}} ,\dots,\tX^d_{\mathrm{k+j-1|k}}  \right), \\ \nonumber
%\mathsf{X}^d_{\mathrm{k+j-l|k}} \in \R^d \  \ \forall \ l,j=1,\dots,N.
%\end{gather}
%Then, the linear models as functions of $\tX^c$ in each leaf of the tree $\mathcal{T}_j$ are defined as
%\begin{gather}
%\label{E:model_leaf}
%\tY_{\mathrm{k+j|k}} =  \beta^T_j [1,\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+j-1|k}} ]^T, \\ \nonumber
%\mathsf{X}^c_{\mathrm{k+j-l|k}} \in \R^c \  \ \forall \ l,j=1,\dots,N.
%\end{gather}
%Note that the coefficients $\beta_j$ would be different for each leaf. Eq.~\eqref{E:model_leaf} implies that the prediction of output $\tY_{\mathrm{k+j}}$ at time $k$ is an affine combination of control inputs from time $k$ to $k+j-1$. Thus, we have managed to linearize the original model dynamics via black-box modeling. This two-step training is done offline. In run-time, given the disturbances $\tX^d_\mathrm{k|k}$ at time $k$, we can narrow down to a leaf of each tree in \eqref{E:model_tree} to retrieve the linear models in \eqref{E:model_leaf}.
%
%In run-time, when a new control action is to be determined, each tree (prediction step) contributes to a linear constraint in the optimization as a replacement for the state dynamics in the case of MPC. Thus, the RHC optimization problem with a quadratic cost ($\mathcal{Q} \geq 0, \mathcal{R} \succeq 0$) can be formulated as:
%\begin{align}
%\begin{aligned}
%\text{min } & \sum_{j=1}^{N} ({\tY}_{\mathrm{k+j|k}})^2 \mathcal{Q} + {\tX^c}^T_{\mathrm{k+j-1|k}} \mathcal{R} {\tX^c}_{\mathrm{k+j-1|k}} +  \lambda\epsilon_j\\
%\text{s.~t. } & \ \ \ \ \ \tY_{\mathrm{k+j|k}} =  \beta^T [1,\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+j-1|k}} ]^T \\
%& \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \underline{\tX}^c \leq \tX^c_{\mathrm{k+j-1|k}} \leq \bar{\tX}^c\\ 
%& \ \ \ \ \ \ \ \ \ \ \ \ \underline{\tY}-\epsilon_j \leq \tY_{\mathrm{k+j|k}} \leq \bar{\tY} + \epsilon_j\\\
%& \ \ \ \ \ \ \ \ \ \ \ \ \ \ \epsilon_j \geq 0, \ j = 1,\dots,N.
%\end{aligned}
%\label{E:dpcrt}
%\end{align}
%Here, $\mathcal{Q} \in \mathbb{R}$ and $\mathcal{R} \in \mathbb{R}^{c \times c}$, and the slack variables $\epsilon_j$ ensure recursive feasibility since the equality constraint on $\tY$ is relaxed. Of course, a different cost function can be chosen depending upon the application. In the current formulation, the data-driven control problem is reduced to a convex program which is much easier to solve than running an optimization directly on a black-box model trained on $\tX^c$ as features. We solve this optimization in the same manner as MPC to determine the optimal sequence of inputs $[\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+N-1|k}}]$, apply the first control input $\tX^c_{\mathrm{k|k}}$ and proceed to the next time step $k+1$. The pseudo code for DPC-RT is given in Algorithm~\ref{A:dpcrt}.
%
%\begin{algorithm}[t!]
%	\caption{Data Predictive Control with Regression Trees}
%	\label{A:dpcrt}
%	\begin{algorithmic}[1]
%		\State \textsc{Design Time}
%		\Procedure{Model Training using Separation of Vars}{}
%		\State Set $\mathsf{X}^c$ $\gets$ manipulated features
%		\State Set $\mathsf{X}^d$ $\gets$ non-manipulated features
%		\State Build $N$ predictive trees with $(\tY,\tX^d)$ defined in \eqref{E:model_tree}
%		\ForAll{trees $\mathcal{T}_j$}
%		\ForAll{regions $R_i$ at the leaves of $\mathcal{T}_j$}
%		\State Fit $ \tY_{\mathrm{k+j|k}} =  \beta^T_j \left[1,\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+j-1|k}} \right]^T$ as in \eqref{E:model_leaf}
%		\EndFor
%		\EndFor
%		\EndProcedure
%		\State \textsc{Run Time}
%		\Procedure{Predictive Control}{}
%		\While{$k< k_{\mathrm{stop}}$}
%		\ForAll{trees $\mathcal{T}_j$}
%		\State Determine the leaf $R_i$ using $\tX^d$ as in \eqref{E:model_tree}
%		\State Obtain the linear model at $R_{i}$ trained in \eqref{E:model_leaf}
%		\EndFor
%		\State Solve optimization in \eqref{E:dpcrt} to determine optimal
%		\State control actions $[\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+N-1|k}}]$
%		\State Apply the first input $\tX^c_{\mathrm{k|k}}$
%		\EndWhile
%		\EndProcedure
%	\end{algorithmic}
%\end{algorithm}
%
%\subsection{DPC-En: DPC with Ensemble Methods}
%\label{SS:dpcrf}
%Regression trees obtain good predictive accuracy in many domains. However, the models used in their leaves have some limitations regarding the kind of functions they are able to approximate.
%The problem with trees is their high variance and that they can overfit the data easily.
%A small change $\delta$ in the data can result in a different series of splits and thus violate the acceptable accuracy $\eps$.
%This is the price to be paid for estimating a tree-based structure from the data.
%
%
%We use ensemble methods \cite{Friedman2001} to combine the predictions of several independent regression trees in order to improve generalizability and robustness over a single estimator. 
%The essential idea is to average many noisy trees to reduce the overall variance in prediction.
%We inject randomness into the tree construction in two ways. First, we randomize the features used to define splitting in each tree.
%Second, we build each tree using a bootstrapped or sub-sampled data set.
%In this way, each tree in the forest is trained on different data, which introduces differences between the trees. More explicitly, training features $\tX^d \in \R^p$ with $p<d$ and the in-bag samples (in-bag samples correspond to the data samples on which the tree was trained) are different for each tree in the forest i.e $|(\tX,\tY)|<n$.
%
%The goal with DPC-En is to replace each tree in Algorithm~\ref{A:dpcrt} by a forest
%\begin{gather}
%\label{E:model_forest}
%\tY_{\mathrm{k+j|k}} = \mathit{f}_{\mathrm{forest}} \left( \tX^d_{\mathrm{k+j-N|k}} ,\dots,\tX^d_{\mathrm{k+j-1|k}}  \right), \\ \nonumber
%\mathsf{X}^d_{\mathrm{k+j-l|k}} \in \R^p \  \ \forall \ l,j=1,\dots,N,
%\end{gather}
%which, again, is trained only on $\mathsf{X}^d$, with $\tX^d \in \R^p \subset \R^d$ for each tree,  and then fit a linear regression model using $\mathsf{X}^c$ in every leaf of every tree. We build $N$ such forests for $N$ prediction steps such that the leaf $R_i$ of forest $\mathcal{R}_j$ uses a linear model
%\begin{gather}
%\label{E:model_leaf_forest}
%\tY_{\mathrm{k+j|k}} =  \Theta^T_{ij} [1,\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+j-1|k}} ]^T, \\ \nonumber
%\mathsf{X}^c_{\mathrm{k+j-l|k}} \in \R^c \  \ \forall \ l,j=1,\dots,N.
%\end{gather}
%Here $(\mathsf{X}^c,\mathsf{Y})$ correspond to the in-bag samples for the trees.
%
%While the offline training burden in DPC-En is slightly increased compared to DPC-RT, in the control step we exploit the better accuracy, and lower variance properties of the random forest. 
%If a forest has $t$ number of trees, given the forecast of disturbances, we have $t$ sets of linear coefficients. We simply average out all the coefficients from all the trees to get one linear model represented by $\hat{\Theta}_j$ for each forest. Note that the averaging step can only be done in run-time because the leaf of each tree can be narrowed down only when the $\tX^d$ is known. Thus, for $N$ forests, we again have exactly $N$ linear equality constraints in the optimization problem below:
%\begin{align}
%\begin{aligned}
%\text{min } & \sum_{j=1}^{N} ({\tY}_{\mathrm{k+j|k}})^2 \mathcal{Q} + {\tX^c}^T_{\mathrm{k+j-1|k}} \mathcal{R} {\tX^c}_{\mathrm{k+j-1|k}} +  \lambda\epsilon_j\\
%\text{s.~t. } & \ \ \ \ \ \tY_{\mathrm{k+j|k}} =  \hat{\Theta}^T_j [1,\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+j-1|k}} ]^T \\
%& \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \underline{\tX}^c \leq \tX^c_{\mathrm{k+j-1|k}} \leq \bar{\tX}^c\\ 
%& \ \ \ \ \ \ \ \ \ \ \ \ \underline{\tY}-\epsilon_j \leq \tY_{\mathrm{k+j|k}} \leq \bar{\tY} + \epsilon_j\\\
%& \ \ \ \ \ \ \ \ \ \ \ \ \ \ \epsilon_j \geq 0, \ j = 1,\dots,N.
%\end{aligned}
%\label{E:dpcrf}
%\end{align}
%DPC-En is graphically described in Figure~\ref{F:dpc-algo-rf}. The ensemble data predictive control (DPC-En) is the first such method to bridge the gap between ensemble predictive models (such as random forests) and receding horizon control. In the next section, we compare DPC-RT and DPC-En to MPC for a building model.
%
%\begin{figure}[t!]
%	\centering
%	\includegraphics[width=26pc]{figures/dpc-algo-rf.eps}
%	\caption{DPC-En: At time $k$, the algorithm uses the forecast of disturbances $\tX^d_{\mathrm{k|k}}$ to select linear models $\Theta_1$ to $\Theta_t$ in the leaves of each ensemble. The linear models in each ensemble are averaged to calculate a single model represented by $\hat{\Theta}_j$ which act as constraints in the optimization problem. The optimal sequence $[\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+N-1|k}}]$, of which the first one is applied, and $\tX^d_{\mathrm{k+1|k+1}}$ is calculated to proceed to $k+1$.}
%	\label{F:dpc-algo-rf}
%\end{figure}