### 2018-01-22 paper/final/cs-challenges.tex

\begin{enumerate}
\item \emph{\textbf{Data Availability:}} \textcolor[rgb]{0,0,1}{The main practical challenge for DPC lies in the availability of data for training, and we require answers to questions like \emph{how much data is required, and how should the sampling (functional testing) be done?} Therefore, the procedure for optimal experiment design, and model improvement with estimation of variance in predictions is one of the main focus of our ongoing work \cite{JainICCPS2018}}.
\item \emph{\textbf{Stability:}} While the buildings are inherently stable, many other applications, such as power networks, require stability guarantees. In our ongoing work, we are working towards proving asymptotic stability to origin with DPC-RT and DPC-En by using concept of switched LTI systems. This will make DPC useful for systems with faster dynamics.
\end{enumerate}