### 2018-01-22 paper/final/Highlights.tex
review of Sec 6 and remark added in Sec3

\documentclass[12pt]{article}

\begin{document}

\title{Highlights}
\date{}
\maketitle 
\thispagestyle{empty}

\begin{itemize}
	% Remember maximum 85 characters (included spaces, points, commas, etc.)
	\item Predictive control strategy using data-driven models via Random Forests.
	\item Data Predictive Control shows comparable performance to Model Predictive Control.
	\item Data Predictive Control shows good performance in Demand Response curtailment.
	\item Data Predictive Control allows energy saving up to $50\%$ in considered case study.
	\item Data Predictive Control is robust to data and weather forecast uncertainties.
\end{itemize}

\end{document}